Denotational Semantics of General Payment Primitives, and Its Payment System

### 2022-09-22 packages/spec-haskell/yellowpaper/Paper.tex
subsection WIP: Payment Execution Environment Models

\begin{itemize}
\item \textit{money distribution} models how monetary value is distributed amongst bearers\footnote{(Banking \& Finance)
a person who presents a note or bill for payment. - Collins English Dictionary},
\item \textit{payment primitives} updates money distribution,
\item \textit{financial system} provides the execution environment for the payment primitives,
\item and \textit{forms of money medium} are the "user interfaces" of money for the bearers.
\end{itemize}

@@

\subsubsection{Haskell Definitions}

@@

Payment primitives are functions that takes a money distribution, shared context and some arbitrary arguments as inputs
and an updated money distribution and shared context as outputs:

\begin{minted}[escapeinside=||,mathescape=true]{haskell}
    somePaymentPrimitive :: MoneyDistribution |$\otimes$| SharedContext
                         -> Args
                         -> MoneyDistribution |$\otimes$| SharedContext
\end{minted}

Loosely speaking, it is considered a primitive, if it can not be broken down into other existing primitives which result
in the same money distribution.

@@

\subsection{Payment System Models}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Discrete Payments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Continuous Payments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{displayquote}
A useful observation about existing money schemes is that they all have some kind of monetary units that are physical or
digital representations of money. Examples are bills, coins, bank accounts, Bitcoin UTXOs, etc.
\end{displayquote}

>>>>

\begin{itemize}
\item \textit{money distribution} models how monetary value is distributed amongst bearers\footnote{(Banking \& Finance)
a person who presents a note or bill for payment. - Collins English Dictionary},
\item \textit{payment primitives} updates money distribution,
\item \textit{payment execution environment} performs payment primitives,
\item and \textit{forms of money medium} are the "user interfaces" of money for the bearers.
\end{itemize}

@@

\subsubsection{Haskell Definition Of Money Distribution}

@@

Payment primitives are functions that takes a money distribution, shared context and some payment instruction as inputs
and an updated money distribution and shared context as outputs:

\begin{minted}[escapeinside=||,mathescape=true]{haskell}
    somePaymentPrimitive :: MoneyDistribution |$\otimes$| SharedContext
                         -> PaymentInstruction
                         -> MoneyDistribution |$\otimes$| SharedContext
\end{minted}

Loosely speaking, it is considered a primitive, if it can not be broken down into other existing primitives which result
in the same money distribution; additionally primitives should be the only constructs in a payment system that can
update money distribution.

@@

\subsection{Payment Execution Environment Models}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The purpose of a payment execution environment is to perform the actual payment primitives taking into account their
conditions, timing and order.

\subsubsection{Composing Financial Contracts}

Inspired by the \textit{technique of composing financial contracts} demonstrated in \cite{peyton2000composing}, we first
define a type class for financial contract:

\input{FinancialContract.lhs.tex}

And let $fcSet$ be the set of all financial contracts in the payment system.

\subsubsection{Sequential Model of Execution Environment}

A sequential execution environment model that can support conditional payment primitives\footnote{Timing is a type of
condition of which current system time is a factor} then can be defined as follows:

\begin{equation}
    \begin{split}
        [\![& \text{FS md fcSet t}]\!] = \sum_{x \in fcSet} ( \\
        & \text{if fcPred md x t} \\
        & \text{then fcExec x t} \\
        & \text{else }\emptyset \\
        )
    \end{split}
\end{equation}

\subsubsection{Deterministic Execution Order}

\subsubsection{Concurrency}

@@

\begin{displayquote}
A useful observation about existing money schemes is that they all have some kind of monetary units that are physical or
digital representations of money. Examples are bills, coins, bank accounts, Bitcoin UTXOs, etc. \cite{buldas2021unifying}
\end{displayquote}
